Centralized Programming for Distributed Control Planes

### 2016-01-28 paper/introduction.tex
updates to intro

One reason for network configuration being onerous is that device configuration languages are low-level and several configuration elements must be kept consistent (e.g., interface addresses) for the device to behave as intended.

Another, more fundamental reason is the semantic mismatch between the intended high-level
policies and the low-level device-by-device configurations.  In particular,
many policies involve network-wide properties---prefer a certain neighbor,
never announce a particular destination externally,
use a particular path only if another fails---but configurations describe the behavior of
individual devices.
%
Operators must manually decompose network-wide policy into device-level behaviors, such that their interactions satisfy the high-level objectives.
%
Ensuring policy-compliance through this manual decomposition is challenging under normal
circumstances and even more so in the face of failures. Some decompositions that work correctly
in failure-free environments, violate key constraints when failures occur.
%
As a result, many network configuration problems are only revealed after failures~\cite{batfish},
sometimes making a bad situation worse.

To reduce configuration errors, many practitioners have a adopted a template-based
approach~\cite{hatch,thwack}, in which common configuration patterns are captured as parameterized templates.  More powerful still are systems like
ConfigAssure~\cite{narain:lisa05,narain+:configassure}, which use SAT solving
and model finding tools to fill in parameters in configurations while
ensuring key correctness constraints are satisfied.  However,
 while templates help ensure certain kinds of consistency across devices, 
they do not bridge the
semantic divide between network policies and device-level configuration,
nor do they do not provide fundamentally different abstractions
from existing configuration languages or new, higher-level reasoning 
principles.

Software-defined networking (SDN) and its abstractions 
are, in part, the research
community's response to the difficulty of maintaining policy
compliance through distributed interaction of individual
devices~\cite{sdn-languages}. Indeed, instead of organizing networks
around a distributed collection of devices that compute forwarding tables through
mutual interactions, the devices are told how to
forward packets by a centralized controller. The controller is responsible for ensuring that the
paths taken are compliant with operator specifications.
%Researchers
%have developed increasingly sophisticated languages that let operators
%specify desirable network paths~\cite{x,y,z} which are then translated
%to forwarding tables at runtime.

The centralized control planes of SDN, however, are not a panacea.
First, while many SDN systems provide effective \emph{intra}-domain routing
abstractions, allowing users to specify paths within their own network,
they fail to provide a coherent means to enable specification of preferred
\emph{inter}-domain routes.  Second, centralized control planes
require careful design and engineering to be robust to failures at scale---one must ensure that all devices can communicate with the controller at all times, even under arbitrary combinations of failure. Even ignoring failures, it is necessary for controller systems to
scale to meet the demands of larger, geographically distributed networks, 
and to react sufficiently quickly
to a changing environment. To address this challenge, some researchers are beginning to 
explore
multi-controller networks, with interacting controllers, thus bringing back distributed
control planes~\cite{mccauley2013extending,onos} and their current programming difficulties.  
%However, academic
%language design and implementation efforts have not kept pace.  For instance, work on many
%experimental SDN languages~\cite{frenetic,flowlog,vericon,merlin,netkat,kinetic,pga} has not yet shown how to implement fault tolerant
%multi-controller systems that support their high-level abstractions efficiently.

@@

To achieve the first goal, we borrowed heavily from the design of
recent high-level SDN languages such as FatTire~\cite{fattire}, 
Merlin~\cite{foster:merlin}, and
NetKAT~\cite{netkat} by using regular expressions to specify paths
through networks.  However, our design also contains several key
differences from existing languages.  The most important difference is semantic:  the paths specified
by programmers can extend from outside the operator's network to inside
the network, across several devices internally, and then out again.  This design
choice allows users to specify preferences about both external and internal 
routes in the same uniform notation.  No SDN programming language provides
similar facilities.  In addition,
we incorporate expression of user preferences in to the algebra
of regular expressions and provide a semantics to our language
in terms of sets of ranked paths.  These preferences indicate possible
fail-over routes:  If too many faults occur, the network may be
disconnected, but as long as any path the user has expressed is
still available, the system guarantees the highest-ranked, available
path will be taken.  We also support network operators by defining a
set of common abbreviations for entering, leaving, and traversing
the user network and we provide abstractions for managing common features of
control plane algorithms such as route aggregation.  Operators use familiar logical operations
to build routes in a modular and compositional manner from these 
primitives.  Our abbreviations are translated in to our regular-expression-based
intermediate representation and compiled from there.

To achieve the second goal, we define program analysis and compilation
algorithms that translate the regular policies in to a graph-based
intermediate language and from there in to BGP.  During the
compilation process, our compiler automatically synthesizes per-device
import and export filters, local preferences, MED attributes, and 
community tags.  We chose BGP for pragmatic reasons:  it is an industry
standard and many data centers use it internally as well as externally.
Indeed, despite the presence of SDN, many networks will continue to
use BGP for the foreseeable future due to investments they have already
made in the infrastructure, the difficulty of transitioning to SDN,
and the scalability and fault-tolerance advantages of a distributed
control plane.  However, using BGP as a back end is not without its 
challenges: not all high-level specifications can be implemented correctly in
the presence of arbitrary failures.  Consequently, we define new
program analysis algorithms to detect specifications that cannot be implemented
and report our findings to the user, so he or she may fix the specification
early rather than waiting until configurations have been deployed.  We also
analyze route aggregation to detect the possibility of failure-induced black-holes.
These algorithms differ from
standard BGP analysis algorithms~\cite{feamster:rcc,batfish} because they operate on 
our graph-based intermediate representation, which operates at a higher level
of abstraction than raw BGP configurations (and which has the pleasant effect of simplifying our 
task).

To evaluate the \sysname{} language and system, we have experimented with specifying policy
for several realistic data center and backbone networks and compared these specifications
with existing operating specifications. We have also
measured the performance of our compiler, demonstrating that our compilation
algorithms scale to topologies with several thousand routers.

In the remainder of the paper, we review BGP~\ref{sec:background} (readers
familiar with BGP may skip this section),
further motivate the need for new languages 
for programming distributed control planes (Section \ref{sec:motivation}),
introduce \sysname{} by example (Section \ref{sec:propane}), 
explain our compiler algorithms (Section \ref{sec:compilation}),
present several lemmas justifying the correctness of our techniques (Section \ref{sec:theory}), describe our implementation (Section \ref{sec:implementation}
and discuss related work (Section \ref{sec:related}).

>>>>

%One reason for network configuration being onerous is that device configuration languages are low-level and several configuration elements must be kept consistent (e.g., interface addresses) for the device to behave as intended.

Besides the languages being low-level, a fundamental reason for configuration being onerous is the semantic mismatch between the intended high-level
policies and the low-level device-by-device configurations.  In particular,
many policies involve network-wide properties---prefer a certain neighbor,
never announce a particular destination externally,
use a particular path only if another fails---but configurations describe the behavior of
individual devices.
%
Operators must manually decompose network-wide policy into device-level behaviors, such that their interactions satisfy the high-level objectives.
%
Ensuring policy-compliance through this manual decomposition is challenging under normal
circumstances and even more so in the face of failures. Some decompositions that work correctly
in failure-free environments, violate key constraints when failures occur.
%
As a result, many network configuration problems are only revealed after failures~\cite{batfish},
sometimes making a bad situation worse.

To reduce configuration errors, many practitioners have a adopted a template-based
approach~\cite{hatch,thwack}, in which common tasks are captured as parameterized templates.
%
%More powerful still are systems like
%ConfigAssure~\cite{narain:lisa05,narain+:configassure}, which use SAT solving
%and model finding tools to fill in parameters in configurations while
%ensuring key correctness constraints are satisfied.
One step further, systems like
ConfigAssure~\cite{narain:lisa05,narain+:configassure} use SAT solving
and model finding tools to correctly and consistently fill in some parameters.
%
However,
 while templates help ensure certain kinds of consistency across devices,
they do not bridge the
semantic divide between network policies and device-level configuration.
They do not provide fundamentally different abstractions
from existing configuration languages or new, higher-level reasoning
principles.

Software-defined networking (SDN) and its abstractions
are, in part, the research
community's response to the difficulty of maintaining policy
compliance through distributed interaction of individual
devices~\cite{ethane}. Indeed, instead of organizing networks
around a distributed collection of devices that compute forwarding tables through
mutual interactions, the devices are told how to
forward packets by a centralized controller. The controller is responsible for ensuring that the
paths taken are compliant with operator specifications.
%Researchers
%have developed increasingly sophisticated languages that let operators
%specify desirable network paths~\cite{x,y,z} which are then translated
%to forwarding tables at runtime.

The centralized control planes of SDN, however, are not a panacea.
First, while many SDN systems provide effective \emph{intra}-domain routing
abstractions, letting users specify paths within their network,
they fail to provide a coherent means to specify preferred
\emph{inter}-domain routes.  Second, centralized control planes
require careful design and engineering to be robust to failures---one must ensure that all devices can communicate with the controller at all times, even under arbitrary combinations of failure. Even ignoring failures, it is necessary for controller systems to
scale to meet the demands of large, geographically distributed networks,
and to react quickly
to a changing environment. To address this challenge, researchers are exploring
multi-controller systems with interacting controllers, thus bringing back distributed
control planes~\cite{mccauley2013extending,onos} and their current programming difficulties.
%However, academic
%language design and implementation efforts have not kept pace.  For instance, work on many
%experimental SDN languages~\cite{frenetic,flowlog,vericon,merlin,netkat,kinetic,pga} has not yet shown how to implement fault tolerant
%multi-controller systems that support their high-level abstractions efficiently.

@@

To achieve the first goal, we borrowed heavily from 
recent high-level SDN languages such as FatTire~\cite{fattire},
Merlin~\cite{foster:merlin}, and
NetKAT~\cite{netkat} by using regular expressions to specify network paths.  However, our design also contains several key
departures from existing languages.  The most important difference is semantic:  the paths specified
by programmers can extend from outside the operator's network to inside
the network, across several devices internally, and then out again.  This design
choice allows users to specify preferences about both external and internal
routes in the same notation.  
%No SDN programming language provides similar facilities.  
In addition,
we incorporate user preferences in to the algebra
of regular expressions and provide a semantics in terms of sets of ranked paths.  The preferences indicate fail-over behaviors:  among all specified paths that are still available,
the system guarantees that the highest-ranked one will be taken.  
%We also support network operators by defining a
%set of common abbreviations for entering, leaving, and traversing
%the user network and we provide abstractions for managing common features %of
%control plane algorithms such as route aggregation.  Operators use familiar logical operations
%to build routes in a modular and compositional manner from these
%primitives.  Our abbreviations are translated in to our regular-expression-based
%intermediate representation and compiled from there.

To achieve the second goal, we develop program analysis and compilation
algorithms that translate the regular policies to a graph-based
intermediate language and from there to BGP configurations.  
We target BGP for pragmatic reasons:  it is an industry
standard and many networks use it internally as well as externally. Despite the advent of SDN, many networks will continue to
use BGP for the foreseeable future due to existing infrastructure investments, the difficulty of transitioning to SDN,
and the scalability and fault-tolerance advantages of a distributed
control plane.

However, using BGP as a vehicle for implementing high-level policies is
challenging because not all policies can be implemented correctly in
the presence of arbitrary failures.  We develop new
algorithms to detect policies that cannot be implemented
and report our findings to the operators, so they may fix the specification at compile time rather than experience undesirable behavior after configurations are deployed.
%We also analyze route aggregation to detect the possibility of failure-induced black-holes.
If the policies are implementable, our compiler automatically synthesizes per-device import and export filters, local preferences, MED attributes, and community tags.

%These algorithms differ from
%standard BGP analysis algorithms~\cite{feamster:rcc,batfish} because they operate on
%our graph-based intermediate representation, which operates at a higher level
%of abstraction than raw BGP configurations (and which has the pleasant effect of simplifying our
%task).

We have implemented our language and compiler in a system called \sysname. To evaluate it, we use it to specify real policies of data center and backbone networks.
% and compared these specifications with existing configurations. 
We find that our language expresses such policies easily, and that the compiler scales to topologies with hundreds of routers.

%measured the performance of our compiler, demonstrating that our compilation
%algorithms scale to topologies with several thousand routers.
%
%\dpw{The following could be cut for space.}
%In the remainder of the paper, we review BGP (Section~\ref{sec:background}---readers
%familiar with BGP may skip this section),
%further motivate the need for new languages
%for programming distributed control planes (Section \ref{sec:motivation}),
%introduce \sysname{} by example (Section \ref{sec:propane}),
%explain our compiler algorithms (Section \ref{sec:compilation}),
%present several lemmas justifying the correctness of our techniques (Section \ref{sec:theory}), describe our implementation (Section \ref{sec:implementation})
%evaluate our work  (Section \ref{sec:evaluation}) and discuss related work (Section \ref{sec:related}).
